Interest Flooding Attacks and countermeasures in Named Data Networking

### 2013-01-14 04a-design-naive.tex
Small reorganization of evaluation section. Some spacing changes
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@915 df1a90cd-247b-4364-bdaf-287a704eb499
My pass on section 4 and 5
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@917 df1a90cd-247b-4364-bdaf-287a704eb499

An obvious and na\"ive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture---flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers to control the inbound Data traffic by controlling the number of outstanding Interests in the network. 
One simple implementation technique  is for an NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding out interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of Data requested that can fully utilize the downstream link (estimated from the number of Interests it forwarded) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens (the pending \emph{Interests Limit}) for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:
%
\begin{equation}
\small \mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

@@

This {\it Token Bucket} approach might be exceptionally retrictive in forwarding Interests---not all Interests will result in a Data packet---thus underutilizing the network. However, the biggest drawback of the algorithm is that it can be used as an effective mechanism to carry out denial of service attacks. If a router has utilized all its tokens to forward malicious Interests, it can no longer forward incoming Interests from legitimate users till the pending malicious Interests start to expire. One way to get around this issue is to impose a per interface limit, so that malicious Interests are not allowed to entirely consume the limits of a specific interface. We describe this technique in greater detail below.

@@

To address the lack of fairness associated with the na\"ive {\it Token Bucket} algorithm, we modify it to ensure that the Interests forwarded by a router on each interface represent a fair mix of Interests received from neighboring interfaces. For example, in Fig.~\ref{fig:flooding example} router A can ensure that the tokens associated with Interests sent out on interface {\texttt eth2}  are fairly distributed across incoming interfaces \texttt{eth0} and \texttt{eth1}. 
%Due to the expected small volume of Interests, we cannot merely rely on network buffers to perform statistical multiplexing of Interests, as they would almost never be %buffered.
%At the same time, until bag of tokens is not empty, there is no reason to delay Interest forwarding, as we do not known how many and from which interfaces Interests %will arrive in the future.
In order to achieve our goal of ensuring ``fair'' mixing of Interests from all neighboring interfaces, 
%, we implement an additional features to support buffering of Interests, if they cannot be immediately forwarded. 
we extend the Pending Interest Table to support flagging of Interests that cannot be immediately forwarded and implement hierarchical queues for each interface (see Fig.~\ref{fig:queueing}). 
This mechanism is essentially a class based queuing~\cite{floyd1995link}, with classes for each outgoing/incoming interface.
%As an alternative to hierarchical queues, one can also use an approach based on virtual time. For details, please see ~\cite{zhang1990virtual}.
We note that unlike normal queuing, Interest queues do not actually store a packet, but merely a bi-directional pointer to the existing PIT entry.
Thus, a PIT entry can be quickly updated when the Interest is actually forwarded; further the element can be quickly removed from the queue when the Interest expires.

@@

\begin{figure}[thb]
  \centering
  \includegraphics[scale=0.65]{queue}
  \vspace{-1cm}
  \caption{Interest queuing: if tokens are unavailable, the router creates a PIT entry, but instead of forwarding, it enqueues the Interest}
  \label{fig:queueing}
\end{figure}

We present a formalized description of this algorithm in Pseudocode~\ref{alg:queuing}. 
By setting appropriate queue sizes we can control the amount of physical resources utilized at a router (such as memory and computation power). 
It is also important to set apprpriate value for how long an Interest can be enqueued. 
If an Interest is enqueued for a long time, by the time it is dequeued and appropriately forwarded, the resulting Data packet might not be valid due to state expiration at the downstream routers. 
For our evaluations we chose to enqueue Interests up to 10\% of their original lifetime.
% We believe that implementing additional mechanisms on each pair of communicating routers to keep the Interests from getting stale would mitigate this issue, though the details are beyond the scope of this paper.
% Alex: I don't believe there is a need for separate discussion about "freshness", as we already introduced concept of lifetime
% {\color{red} \it Priya:  We need to  touch upon the Interest freshness issue in the NDN overview section; else we should get rid of this bit.}

@@

\floatname{algorithm}{Pseudocode}

@@

As we show in Section~\ref{sec:evaluation}, this algorithm provides partial relief from the Interest flooding attacks, allowing legitimate users to successfully fetch Data for 15--20\% of their expressed Interests. 
We note that while this algorithm might be reasonable for ensuring fairness in an NDN network, it is ineffective in protecting legitimate users from malicious attackers. 
Malicious attackers are able to successfully thwart access to content for legitimate users by sending a relatively modest volume of Interests.

@@

The key drawback of the {\it Token bucket with per interface fairness} algorithm is that it admits Interests from malicious users. A sizeable percentage of these malicious Interests are forwarded all the way to content producers, thereby utilizing scarce network resources and reducing resources available to serve legitimate users.  This algorithm attempts to ensure that each interface does not forward more than its fair share of Interests, but in doing so, it drops both legitimate Interests as well as malicious Interests. For any strategy to be effective in defending against Interest flooding attacks, it must be able to detect and effectively drop only malicious requests for content. `Good' Interests from legitimate users must be admited and forwarded appropriately. The key question is how can we devise mitigation algorithms that allow a router to distinguish between `good' and `bad' Interests?

@@

In NDN one cannot easily distinguish a `good' prefix from a `bad' one. Content producers must be able to support dynamic generation of Data packets to satisfy incoming Interest packets---thus, an Interest that cannot be satisfied by matching Data from any intermediate router's cache is necessarily forwarded to the producer for possible dynamic Data packet generation. 
Thus intermediate routers cannot determine if the prefix is `good' or `bad' based on the prefix name alone---DDoS mitigation techniques that rely on whitelisting or blacklisting of certain namespaces  will be inefficient and likely ineffective.

>>>>

An obvious and na\"ive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture---flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers to control the inbound data traffic by controlling the number of outstanding Interests in the network. 
One simple implementation technique  is for an NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of data requested that can fully utilize the downstream link (estimated from the number of forwarded Interests) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens (the pending \emph{Interests Limit}) for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:
%
\begin{equation}
\small \mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

@@

This {\it Token Bucket} approach might be exceptionally retrictive in forwarding Interests---not all Interests will result in a Data packet---thus, the network can be underutilized. However, the biggest drawback of the algorithm is that it itself can be used as an effective mechanism to carry out DDoS attacks. If a router has utilized all its tokens to forward malicious Interests, it can no longer forward incoming Interests from legitimate users till the pending malicious Interests start to expire. One way to get around this issue is to impose a per interface limit, so that malicious Interests are not allowed to entirely consume the limits of a specific interface. We describe this technique in greater detail below.

@@

To address the lack of fairness associated with the na\"ive {\it Token Bucket} algorithm, we modify it to ensure that the Interests forwarded by a router on each interface represent a fair mix of Interests received from neighboring nodes. For example, in Fig.~\ref{fig:flooding example} router A can ensure that the tokens associated with Interests sent out on interface {\texttt eth2}  are fairly distributed across incoming interfaces \texttt{eth0} and \texttt{eth1}. 
%Due to the expected small volume of Interests, we cannot merely rely on network buffers to perform statistical multiplexing of Interests, as they would almost never be %buffered.
%At the same time, until bag of tokens is not empty, there is no reason to delay Interest forwarding, as we do not known how many and from which interfaces Interests %will arrive in the future.
In order to achieve our goal of ensuring ``fair'' mixing of Interests from all neighboring nodes, 
%, we implement an additional features to support buffering of Interests, if they cannot be immediately forwarded. 
we extend the Pending Interest Table to support flagging of Interests that cannot be immediately forwarded and implement hierarchical queues for each interface (see Fig.~\ref{fig:queueing}). 
This mechanism is essentially a class based queuing~\cite{floyd1995link}, with classes for each outgoing and incoming interface.
%As an alternative to hierarchical queues, one can also use an approach based on virtual time. For details, please see ~\cite{zhang1990virtual}.
We note that unlike normal queuing, Interest queues do not actually store a packet, but merely a bi-directional pointer to the existing PIT entry.
Thus, a PIT entry can be quickly updated when the Interest is actually forwarded; further the element can be quickly removed from the queue when the Interest expires.

@@

\begin{figure}[thb]
  \centering
  \includegraphics[scale=0.65]{queue}
  \vspace{-0.7cm}
  \caption{Interest queuing: if tokens are unavailable, the router creates a PIT entry, but instead of forwarding, it enqueues the Interest}
  \label{fig:queueing}
\end{figure}

We present a formalized description of this algorithm in Pseudocode~\ref{alg:queuing}. 
By setting appropriate queue sizes we can control the amount of physical resources utilized at a router.
 % (such as memory and computation power). 
It is also important to set appropriate value for how long an Interest can be enqueued. 
If an Interest is enqueued for a long time, by the time it is dequeued and forwarded, the retrieved Data packet could be dropped later due to state expiration at the downstream routers. 
For our evaluations we chose to enqueue Interests up to 10\% of their original lifetime.
% We believe that implementing additional mechanisms on each pair of communicating routers to keep the Interests from getting stale would mitigate this issue, though the details are beyond the scope of this paper.
% Alex: I don't believe there is a need for separate discussion about "freshness", as we already introduced concept of lifetime
% {\color{red} \it Priya:  We need to  touch upon the Interest freshness issue in the NDN overview section; else we should get rid of this bit.}

@@

\floatname{algorithm}{\small Pseudocode}

@@

As we show in Section~\ref{sec:evaluation}, this algorithm provides partial relief from the Interest flooding attacks, allowing legitimate users to successfully fetch Data for 15--20\% of their expressed Interests. 
We note that while this algorithm might be reasonable for ensuring limited fairness in an NDN network, it is ineffective in protecting legitimate users from malicious attackers. 
Malicious attackers are able to successfully thwart access to content for legitimate users by sending a relatively modest volume of Interests.

@@

The key drawback of the {\it Token bucket with per interface fairness} algorithm is that it admits Interests from malicious users. A sizeable percentage of these malicious Interests are forwarded all the way to content producers, thereby utilizing scarce network resources and reducing resources available to serve legitimate users.  This algorithm attempts to ensure that each interface does not forward more than its fair share of Interests, but in doing so, it drops both legitimate Interests as well as malicious Interests. For any strategy to be effective in defending against Interest flooding attacks, it must be able to detect and effectively drop only malicious requests for content. 
% `Good' Interests from legitimate users must be admited and forwarded appropriately. 
The key question is how can we devise mitigation algorithms that allow a router to distinguish between `good' and `bad' Interests?

@@

{\color{red} \textbf{Alex: This paragraph is confusing to me. I would remove it.} In NDN one cannot easily distinguish a `good' prefix from a `bad' one. Content producers must be able to support dynamic generation of Data packets to satisfy incoming Interest packets---thus, an Interest that cannot be satisfied by matching Data from any intermediate router's cache is necessarily forwarded to the producer for possible dynamic Data packet generation. 
Thus intermediate routers cannot determine if the prefix is `good' or `bad' based on the prefix name alone---DDoS mitigation techniques that rely on whitelisting or blacklisting of certain namespaces  will be inefficient and likely ineffective. }

### 2013-01-14 04b-design-intelligent.tex
Small reorganization of evaluation section. Some spacing changes
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@915 df1a90cd-247b-4364-bdaf-287a704eb499
Small spacing changes (there was text overlapping on figures)
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@916 df1a90cd-247b-4364-bdaf-287a704eb499
My pass on section 4 and 5
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@917 df1a90cd-247b-4364-bdaf-287a704eb499

In order to distinguish between legitimate and malicious Interests, we leverage another unique feature of  NDN architecture---guaranteed symmetric flow of Interest and Data packets. Since a Data packet takes the reverse path of the corresponding Interest packet, a router is guarateed to see if an Interest it forwarded resulted in a matching Data packet. The only exception being if Interest/Data packets were lost along the way due to congestion in the network.
Therefore, intermediate routers can classify all Interests that result in matching Data as legitimate, while the ones that timed out can be marked as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest flooding}).}  
%Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

@@

Fig.~\ref{fig:ratio example} illustrates the resulting dynamics of the statistic during and after an Interest flooding attack. The attack duration is from 10 to 70~seconds. Prior to start of the attack, the percentage of unsatisfied Interests is zero.  
The statistics build up rapidly as soon as Interests start to time out, which happens approximately one second after the start of the attack.\footnote{Again, we are assuming that Interests are admitted for a maximum period one second.}
For the duration of the attack (10--70~seconds), the percentage of unsatisfied Interests is close to 100\%: 
when the ratio is close to 100\%, routers drop all incoming Interests, resulting in decaying of the statistics until a new Interest is admitted, which eventually brings statistics back near 100\% point.
Finally, the ratio exponentially decays after the attack ceases.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.8]{limits}
  \caption{Dynamics of the unsatisfied Interests statistics on gateway's interface towards the attacker}
  \label{fig:ratio example}
\end{figure}

@@

A drawback of the satisfaction-based Interest acceptance method is that each router on the path makes an independent decision on whether to forward or drop the Interest. 
As a result of these independent decisions,  the probability of legitimate Interests being forwarded decreases rapidly as the number of hops between the content requester and producer grows; worsening the Interest satisfaction statistics and resulting in further drops.
In example on Fig.~\ref{fig:flooding example}, the router A observes 50\% satisfaction rate for \texttt{eth1} and 0\% rate for \texttt{eth0}. 
At the same time, router B observes a 30\% satisfaction rate for its \texttt{eth0} interface.
Next time a legitimate Interest arrives at router A, it has a 50\% chance of being forwarded further, and if forwarded, it has only a $50\% \times 30\% = 15\%$ probability of being forwarded further towards the Data producer. With each increasing hop in the network, the probability of being forwarded to the next hop decreases significantly. 
One way to prevent this overreaction and unfair penalization is to ensure that the decision taken at each router on whether to forward or drop the Interest is not independent of the decision taken at preceding routers. An explicit notification such as a gossip protocol between neighboring NDN routers that specifies the volume of Interests each router is willing to forward will likely address this issue. We leave the implementation and evaluation of a gossip protocol to future work.
%{\color{red}Alex: we should state that it is out of scope of the paper to evaluate this issue}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{dynamic-limits}
  \caption{Satisfaction-based pushback example
%: routers explicitly tell neighbors how many Interest packets they can deliver to the Data producer%
}
  \label{fig:dynamic limits example}
\end{figure}

\floatname{algorithm}{Pseudocode}

@@

Both satisfaction-based Interest acceptance and satisfaction-based pushback algorithms are forms of a well-known push-back mechanism~\cite{Pushback}, but with several core differences. 
First, we are suppressing (pushing back) unwanted requests for Data, not actual Data itself.
Second, differentiating between good and bad Interests is based on the traffic symmetry principle of NDN.
% Alex: I'm not entirely sure about this point... 
Finally, both intelligent attack mitigation algorithms can be deployed at all times without degrading network performance even when there are no active attackers.

>>>>

In order to distinguish between legitimate and malicious Interests, we leverage another unique feature of  NDN architecture---guaranteed symmetric flow of Interest and Data packets. Since a Data packet takes the reverse path of the corresponding Interest packet, a router is guarateed to see if an Interest it forwarded resulted in a matching Data packet. The only exception being if Interest/Data packets were lost along the way due to congestion in the network.
Therefore, intermediate routers can classify all Interests that result in matching Data as legitimate, while the ones that timed out can be marked as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest-flooding}).}  
%Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

@@

Fig.~\ref{fig:ratio example} illustrates the resulting dynamics of the statistic during and after an Interest flooding attack. The attack duration is from 10 to 70~seconds. Prior to start of the attack, the percentage of unsatisfied Interests is zero.  
The statistics build up rapidly as soon as Interests start to time out, which happens approximately one second after the start of the attack.%\footnote{Again, we are assuming that Interests are admitted for a maximum period one second.}
For the duration of the attack (10--70~seconds), the percentage of unsatisfied Interests is close to 100\%: 
when the ratio is close to 100\%, routers drop all incoming Interests, resulting in decaying of the statistics until a new Interest is admitted, which eventually brings statistics back near 100\% point.
Finally, the ratio exponentially decays after the attack ceases.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.8]{limits}
  \vspace{-0.3cm}
  \caption{Dynamics of the unsatisfied Interests statistics on gateway's interface towards the attacker}
  \label{fig:ratio example}
\end{figure}

@@

A drawback of the satisfaction-based Interest acceptance method is that each router on the path makes an independent decision on whether to forward or drop the Interest. 
As a result of these independent decisions,  the probability of legitimate Interests being forwarded decreases rapidly as the number of hops between the content requester and producer grows; worsening the Interest satisfaction statistics and resulting in further drops.
In example on Fig.~\ref{fig:flooding example}, the router A observes 50\% satisfaction rate for \texttt{eth1} and 0\% rate for \texttt{eth0}. 
At the same time, router B observes a 30\% satisfaction rate for its \texttt{eth0} interface.
Next time a legitimate Interest arrives at router A, it has a 50\% chance of being forwarded further, and if forwarded, it has only a $50\% \times 30\% = 15\%$ probability of being forwarded further towards the data producer. With each increasing hop in the network, the probability of being forwarded to the next hop decreases significantly. 
One way to prevent this overreaction and unfair penalization is to ensure that the decision taken at each router on whether to forward or drop the Interest is not independent of the decision taken at preceding routers. An explicit notification such as a gossip protocol between neighboring NDN routers that specifies the volume of Interests each router is willing to forward will likely address this issue. We leave the implementation and evaluation of a gossip protocol to future work.
%{\color{red}Alex: we should state that it is out of scope of the paper to evaluate this issue}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{dynamic-limits}
  \vspace{-0.3cm}
  \caption{Satisfaction-based pushback example
%: routers explicitly tell neighbors how many Interest packets they can deliver to the Data producer%
}
  \label{fig:dynamic limits example}
\end{figure}

\floatname{algorithm}{\small Pseudocode}

@@

Both satisfaction-based Interest acceptance and satisfaction-based pushback algorithms are forms of a well-known push-back mechanism~\cite{Pushback}, but with several core differences. 
First, we are suppressing (pushing back) unwanted requests for data, not actual data itself.
Second, differentiating between good and bad Interests is based on the traffic symmetry principle of NDN.
% Alex: I'm not entirely sure about this point... 
Finally, both intelligent attack mitigation algorithms can be deployed at all times without degrading network performance even when there are no active attackers.

### 2013-01-14 05-evaluation.tex
My pass on section 4 and 5
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@917 df1a90cd-247b-4364-bdaf-287a704eb499

In this section, we present an in-depth evaluation study, aiming to quantify the effectiveness of all our Interest flooding attack mitigation methods.
We use the open-source ndnSIM~\cite{ndnsim} package, which implements NDN protocol stack for NS-3 network simulator\footnote{\url{http://www.nsnam.org/}}, to run simulations for a variety of network topologies and scenarios. We implement our three mitigation algorithms---token bucket with per interface fairness, satisfaction-based Interest acceptance, or satisfaction-based pushback---in ndnSIM. For evaluating the effectiveness of each mitigation algorithm, every router in the simulated topology runs the mitigation algorithm under study.

@@

The metric we choose to quantify the effectiveness of our defense strategies is the {\it percentage of satisfied Interests for legitimate users}. This metric corresponds to the quality of service experienced by legitimate users when the network is under attack. In other words, if the network implements a mitigation method $X$ and a high percentage of user-expressed Interests are satisfied even while the network is under attack, then one can conlcude that method $X$ is highly effective at mitigating the attack. We also ensure that all Interests expressed by legitimate users during a period of no-attack are satisfied.

@@

We ran our simulations on two different network topologies---a smaller binary tree topology and a much larger ISP topology. We use a binary tree topology as it represents one of the worst cases to defend against Interest flooding DDoS attacks. The larger ISP topology reflects how our mitigation methods would perform when deployed on the real Internet.
Again, to study the performance of our mitigation strategies under a range of conditions, we also varied the percentage of attackers in the network - the values ranged from 6\% attackers to over 50\% attackers in the network.

@@

% Topology description
%To assess baseline quality of the designed Interest flooding attack mitigation methods, we evaluated them first using a %simplistic small-scale binary tree topology 
In Fig.~\ref{fig:small-scale}, we depict the binary-tree topology that we use for our initial experiments.
Legitimate users as well as attackers were placed on leaf nodes (red nodes) as show in the figure. There are 16 end users (both legitimate and attackers) in this topology, each expressing Interests that are routed towards a single data producer placed at the root of the tree.  Each link in this topology is assigned a bandwidth of 10~Mbps and a randomized propagation delay ranging from 1 to 10 ms.

@@

Our goal is to compare the effectiveness of each mitigation method and quantify the percentage of Interests satisfied for all legitimate users while the network is under attack. For each mitigation algorithm, we perform ten independent simulation runs, where we randomly choose 7 client nodes to represent adversaries while the remaining 9 client nodes represent legitimate users. In each run we simulate a 10-minute attack window (total simulation time was 30 minutes, with attack starting at the 10-minute mark). We plot the minimum and maximum range for observed Interest satisfaction percentages for all legitimate users aggregated across the 10 simulation runs as a function of time for each mitigation algorithm in Fig.~\ref{fig:small-scale attack progress}. Token bucket with per-interface fair queuing performs the worst, while satisfaction-based pushback performs the best, with almost a 100\%  satisfied Interests for all legitimate users.

@@

We observe a successful denial of service attack, where 40\% attackers succeed in significantly shutting down the remaining 60\% legitimate users - a mere 15\% of their Interests are satisfied by Data from the producer. Contrary to expectations, the 60\% legitimate users do not receive at least 60\% of network resources. As described in the previous section, the key limitation of this algorithm is that it still admits Interests from attackers, a good percentage of which traverse all the way to the data producer. Significantly reducing the number of Interests accepted from attackers is fundamental for improving the effectiveness of any mitigation algorithm.

@@

Our next goal is to study the effectiveness of our mitigation algorithms as a function of increasing adversaries in the network.
To this end, we vary the percentage of attackers in the topology from 6\% to over 50\%. Since the total number of end users in the topology is constant, as the number of attackers increases, the number of legitimate users decreases. All other parameters and experimental set-up are similar to the previous experiment. As before, for each mitigation algorithm, we perform ten independent simulation runs.
%The second set of conducted experiments was aimed to answer the question of the effect and quality of the Interest flooding attack mitigation algorithm under different attack volume.
%To do this, for each algorithm we varied the number of adversary nodes in the topology, keeping the total number of client nodes constant: 1 attacker and 15 legitimate users, 3 attackers and 13 legitimate users, etc.

@@

In Fig.\ref{fig:small-scale-topo boxplot} we present  the Interest satisfaction ratio for legitimate users aggregated over the 10 simulation runs for the first minute of the attack. The results are as expected---for all three mitigation algorithms, as the percentage of attackers in the network increases, the lower is the Interest satisfaction ratio for legitimate users.
In the case of token bucket with per-interface fairness algorithm, just 3 attackers can halve the quality of service for the remaining 13 legitimate users. While both intelligent attack mitigation algorithms also show a decline in service quality as the percentage of attackers increases, this decline is much more gradual. In the case of satisfaction-based pushback algorithm, over 90\% of Interests from legitimate users are satisfied even when over 50\% of the users are malicious.

@@

Next, our goal is to understand the behavior of our mitigation strategies when deployed on a large-scale real network topology. Our ISP topology is based on a modified version of Rocketfuel's AT\&T topology~\cite{rocketfuel}.
%In order to approximate the general structure of the Internet,
% (scale-free structure, customer-provider, and peer-to-peer relations)
We extract the largest connected component comprising of 562 nodes from this original topology and separate the nodes into three categories: clients, gateways, and backbones. Nodes having degree than four are classified as clients (344 red nodes as shown in Fig.~\ref{fig:large-scale}), nodes directly connected to clients are classified as gateways (109 green nodes), and the remaining nodes as classified as backbones (109 blue nodes). 
%(To ensure that paths in the topology are ``valley-free,'' we augmented the topology with necessary backbone-to-backbone links.) 
We assign bandwidth and  delay values to links based on their type---both values are random numbers within the respective ranges as shown in Table~\ref{tab:large-scale}. We experiment with placing the data producer at both a gateway node as well as backbone node, which we randomly pick for each simulation run. Similar to the binary tree topology experiments, we fix the number of malicious nodes at approximately 40\% (140 out of 344 client nodes in the topology) and randomly pick these nodes for each simulation run. We conduct 10 simulation runs for each mitigation algorithm, with the attack duration spanning a 5 minute interval.

\begin{figure}[htbp]
  \centering
  \vspace{-.1cm}\includegraphics[scale=0.15,angle=90]{7018-r0}
  \caption{Internet-like topology: 344 client routers (red), 109 gateway routers (green), 109 backbone routers (blue)}\vspace{-.2cm}
  \label{fig:large-scale-topo}
\end{figure}

@@

In Fig.~\ref{fig:large-scale}, we summarize our results aggregated over 10 simulations runs for each mitigation algorithm for the scenario where the data producer is placed at a gateway node. We observe similar results for the data producer placed at the backbone node as well. As in the case of the binary-tree topology experiments, token bucket with per interface fairness is the most ineffective algorithm and satisfaction-based pushback is the most effective one.   Interest satisfaction percentage for legitimate users are approximately 25\% and almost 100\% respectively for these two mitigation methodologies.

>>>>

In this section, we present an in-depth evaluation study, aiming to quantify the effectiveness of all our Interest flooding attack mitigation methods.
We used the open-source ndnSIM~\cite{ndnsim} package, which implements NDN protocol stack for NS-3 network simulator (\url{http://www.nsnam.org/}), to run simulations for a variety of network topologies and scenarios. 
We extended ndnSIM with our three mitigation algorithms---token bucket with per interface fairness, satisfaction-based Interest acceptance, or satisfaction-based pushback---and evaluated effectiveness of each algorithm independently.

% For evaluating the effectiveness of each mitigation algorithm, every router in the simulated topology runs the mitigation algorithm under study.

@@

The metric we choose to quantify the effectiveness of our algorithms is the {\it percentage of satisfied Interests for legitimate users}. This metric corresponds to the quality of service experienced by legitimate users when the network is under attack. In other words, if the network implements a mitigation method $X$ and a high percentage of user-expressed Interests are satisfied even while the network is under attack, then one can conlcude that method $X$ is highly effective at mitigating the attack. We also ensure that all Interests expressed by legitimate users during a period of no-attack are satisfied.

@@

We ran our simulations on two different network topologies---a smaller binary tree topology and a much larger ISP-like topology. We use a binary tree topology as it represents one of the worst cases to defend against Interest flooding DDoS attacks. The larger ISP topology reflects how our mitigation methods would perform when deployed on the real Internet.
Again, to study the performance of our mitigation strategies under a range of conditions, we also varied the percentage of attackers in the network---the values ranged from 6\% attackers to over 50\% attackers in the network.

@@

% Topology description
%To assess baseline quality of the designed Interest flooding attack mitigation methods, we evaluated them first using a %simplistic small-scale binary tree topology 
In Fig.~\ref{fig:small-scale}, we depict the binary-tree topology that we used for our initial experiments.
Legitimate users as well as attackers were placed on leaf nodes (red nodes) as show in the figure. There are 16 end users (both legitimate and attackers) in this topology, each expressing Interests that are routed towards a single data producer placed at the root of the tree.  Each link in this topology is assigned a bandwidth of 10~Mbps and a randomized propagation delay ranging from 1 to 10 ms.

@@

Our goal is to compare the effectiveness of each mitigation method and quantify the percentage of Interests satisfied for all legitimate users while the network is under attack. For each mitigation algorithm, we perform 10 independent simulation runs, where we randomly choose 7 client nodes to represent adversaries while the remaining 9 client nodes represent legitimate users. In each run we simulate a 10-minute attack window (total simulation time was 30 minutes, with attack starting at the 10-minute mark). We plot the minimum and maximum range for observed Interest satisfaction percentages for all legitimate users aggregated across the 10 simulation runs as a function of time for each mitigation algorithm in Fig.~\ref{fig:small-scale attack progress}. Token bucket with per-interface fair queuing performs the worst, while satisfaction-based pushback performs the best, with almost a 100\%  satisfied Interests for all legitimate users.

@@

We observe a successful DDoS attack, where 40\% attackers succeed in significantly shutting down the remaining 60\% legitimate users---a mere 15\% of their Interests are satisfied by Data from the producer. Contrary to expectations, the 60\% legitimate users do not receive at least 60\% of network resources. As described in the previous section, the key limitation of this algorithm is that it still admits Interests from attackers, a good percentage of which traverse all the way to the data producer. Significantly reducing the number of accepted malicious Interests is fundamental for improving the effectiveness of any mitigation algorithm.

@@

Our next goal is to study the effectiveness of our mitigation algorithms as a function of increasing adversaries in the network.
To this end, we vary the percentage of attackers in the topology from 6\% to over 50\%. Since the total number of end users in the topology is constant, as the number of attackers increases, the number of legitimate users decreases. All other parameters and experimental set-up are similar to the previous experiment. As before, for each mitigation algorithm, we perform 10 independent simulation runs.
%The second set of conducted experiments was aimed to answer the question of the effect and quality of the Interest flooding attack mitigation algorithm under different attack volume.
%To do this, for each algorithm we varied the number of adversary nodes in the topology, keeping the total number of client nodes constant: 1 attacker and 15 legitimate users, 3 attackers and 13 legitimate users, etc.

@@

In Fig.\ref{fig:small-scale-topo boxplot} we present  the Interest satisfaction ratio for legitimate users aggregated over the 10 simulation runs for the first minute of the attack. The results are as expected---for all three mitigation algorithms, as the percentage of attackers in the network increases, the lower is the Interest satisfaction ratio for legitimate users.
In the case of token bucket with per-interface fairness algorithm, just 3 attackers can halve the quality of service for the remaining 13 legitimate users. While both intelligent attack mitigation algorithms also show a decline in service quality as the percentage of attackers increases, this decline is much more gradual and marginal. In the case of satisfaction-based pushback algorithm, over 90\% of Interests from legitimate users are satisfied even when over 50\% of the users are malicious.

@@

Next, our goal is to understand the behavior of our mitigation strategies when deployed on a large-scale real network topology. Our ISP topology is based on a modified version of Rocketfuel's AT\&T topology~\cite{rocketfuel}.
%In order to approximate the general structure of the Internet,
% (scale-free structure, customer-provider, and peer-to-peer relations)
We extract the largest connected component comprising of 562 nodes from this original topology and separate the nodes into three categories: clients, gateways, and backbones. Nodes having degree less than four are classified as clients (344 red nodes as shown in Fig.~\ref{fig:large-scale}), nodes directly connected to clients are classified as gateways (109 green nodes), and the remaining nodes as classified as backbones (109 blue nodes). 
%(To ensure that paths in the topology are ``valley-free,'' we augmented the topology with necessary backbone-to-backbone links.) 
We assign bandwidth and  delay values to links based on their type---both values are random numbers within the respective ranges as shown in Table~\ref{tab:large-scale}. We experiment with placing the data producer at both a gateway node as well as backbone node, which we randomly pick for each simulation run. Similar to the binary tree topology experiments, we fix the number of malicious nodes at approximately 40\% (140 out of 344 client nodes in the topology) and randomly pick these nodes for each simulation run. We conduct 10 simulation runs for each mitigation algorithm, with the attack duration spanning a 5-minute interval.

\begin{figure}[htbp]
  \centering
  \vspace{-.1cm}\includegraphics[scale=0.15,angle=90,height=3.5cm,width=8cm]{7018-r0}
  \caption{Internet-like topology: 344 client routers (red), 109 gateway routers (green), 109 backbone routers (blue)}\vspace{-.2cm}
  \label{fig:large-scale-topo}
\end{figure}

@@

In Fig.~\ref{fig:large-scale}, we summarize our results aggregated over 10 simulations runs for each mitigation algorithm for the scenario, where the data producer is placed at a gateway node. We observe similar results for the data producer placed at the backbone node as well. As in the case of the binary-tree topology experiments, token bucket with per interface fairness is the most ineffective algorithm and satisfaction-based pushback is the most effective one.   Interest satisfaction percentage for legitimate users are approximately 25\% and almost 100\% respectively for these two mitigation methodologies.
